Predicting Solid State Material Platforms for Quantum Technologies

### 2022-01-17 paper_v2.tex
Updates from Overleaf

The most striking difference is found for the empirical approach. The entries labeled as suitable candidates do not exhibit any magnetization, even though the materials crystallize in both polar and non-polar space groups --- contrarily to the two Ferrenti-based approaches. Moreover, we observe that the range of covalent radius and maximum ionic character span a substantially smaller parameter space in the case of the empirical approach than for the two former ones.  
Here, we can begin to deduce some trends, if the findings from the empirical approach are taken as guidance. Where the polarity of the material's space group may be less important than assumed in the Ferrenti based approaches, ionic character and covalent range are distinguished from the other features visualized in Fig.~\ref{fig:parallel-coordinates-approaches}. 
Identification of the features that promote quantum compatible  characteristics could be crucial for targeted quantum device design based on careful material selection, nanostructuring and defect engineering.

\begin{figure}[t] %!htp
    \centering
    \begin{subfigure}{0.7\textwidth}
        \centering
          \input{figures/parallel-coordinates-plots/01-ferrenti-approach-v2.pgf}
    \end{subfigure}
    \begin{subfigure}{0.7\textwidth}
        \centering
          \input{figures/parallel-coordinates-plots/02-augmented-ferrenti-approach-v2.pgf}
    \end{subfigure}
    \begin{subfigure}{0.7\textwidth}
        \centering
          \input{figures/parallel-coordinates-plots/03-insightful-approach-v2.pgf}
    \end{subfigure}
    %\vspace*{10mm}
    \caption{Parallel coordinate plots for the different approaches. To limit the data cluttering, we have randomly collected up to $250$ entries for each class and made the lines transparent. For the empirical approach, we have used all the $187$ suitable candidates. The axes show total magnetization (mag) taken from the Materials Project (MP) database along with the space group (SG), ionic character (ionic char), covalent range (cov range) as calculated from elemental properties, number of elements (num elements) and energy gap (Eg) extracted from MP.} 
    \label{fig:parallel-coordinates-approaches}
\end{figure}

\begin{figure}%[h] %!tbp
    \centering
    \begin{subfigure}{0.3\textwidth}
        \centering
        \includegraphics[width=1\textwidth]{figures/pca-2d-plots/01-ferrenti-approach-v3.pdf}
    \end{subfigure}%
    \begin{subfigure}{0.3\textwidth}
        \centering
        \includegraphics[width=1\textwidth]{figures/pca-2d-plots/02-augmented-ferrenti-approach-v2.pdf}
    \end{subfigure}

    \begin{subfigure}{0.4\textwidth}
        \centering
        \includegraphics[width=1\textwidth]{figures/pca-2d-plots/03-insightful-approach-v3.pdf}
    \end{subfigure}
    %\vspace*{-95mm}
    \caption{Two-dimensional scatter plots for the three different approaches. We have identified the two eigenvectors corresponding to the two largest eigenvalues of the covariance-matrix, that is, the two most important principal components (PC1 and PC2) of the initial data from the Materials Project query. Then, we have transformed the three training sets resulting from the three approaches and visualized them as scatter plots. Green squares display suitable candidates, along with the black (SiC), blue (diamond) and yellow (Si) symbols in the bottom panel, and red triangles represent unsuitable candidates.}
    \label{fig:2dscatterplotpca}
\end{figure}

%MB !! Keep this here or move below??  !!
To further illustrate the complexity of the training sets, we have identified the two largest eigenvalues of the covariance matrix \cite{Hastie2009} of the initial data set from the Materials Project, and transformed the training sets according to the corresponding two eigenvectors. The resulting scatter plots are depicted in Figure~\ref{fig:2dscatterplotpca}. 
Green squares and red triangles represent suitable and unsuitable candidates, respectively. Due to the complexity of reducing the large amount of features down to only two, both suitable and unsuitable candidates for the Ferrenti and the extended Ferrenti approaches are largely overlapping. 
The logic behind categorizing materials in  two classes (suitable and unsuitable) appears to not translate into a distinct separation, at least not in the representation of Fig.~\ref{fig:2dscatterplotpca} for the Ferrenti and extended Ferrenti approaches.  
Hence, using these approaches for predicting QT material hosts could prove challenging for any model that would try to glean a clear-cut boundary between materials that are and are not suitable for QT. 
%Nonetheless, we note that the unsuitable candidates appear to be restricted to a certain parameter space while the suitable candidates extend much further.  
We expect therefore that the two Ferrenti approaches could need supplementary dimensions for further distinguishing between the materials in the two categories (suitable and unsuitable), or that it may prove challenging to find a generalized model based on the training sets that were defined according to the above discussion.

In the case of the empirical approach, on the other hand, we can discern a trend where the upper left part of the figure is dominated by suitable candidates (green, blue, yellow and black symbols). 
Interestingly, we observe that different configurations of the famously quantum compatible materials of diamond (blue symbols), silicon carbide (black symbols) and silicon (yellow symbols) are grouped together but each material class is separated in its own region. 
The unsuitable candidates (red squares) are similarly restricted to the lower right corner and there appears to be a clear barrier between the two (suitable and unsuitable), albeit with some exceptions. This indicates that the empirical approach may be capable of separating materials based on their underlying properties. Thus, devising a logical framework for the data mining process is likely important in order to obtain a suitable training set for the machine learning methods.

@@

%summary - do we need this? 
In summary, using various data extraction tools for a number of materials science databases, we have obtained a data set consisting of over $25.000$ materials. The information from this data set has then been converted into nearly $5000$ physics-informed features.  
After these steps with fine-tuning and selection of entries for our training sets we are now ready to apply the machine learning methods on the test set in order to classify materials as potential candidates for quantum technology applications.

@@

Four well-tested machine learning algorithms were employed in this work: logistic regression, decision trees, and ensemble methods like random forests and gradient boosting \cite{Mehta2019,Hastie2009}. 
These four machine learning methods were applied to the training sets that were optimized using each of the three approaches (Ferrenti, extended Ferrenti and empirical) outlined above. This yielded twelve sets of materials that were predicted as suitable material hosts for QT applications. 
In the evaluation of the machine learning models for the three different approaches we apply a $5\times 5$ stratified cross-validation when searching for the optimal hyperparameter  combinations. 
\textbf{Principal components were found by ...} 
The extracted principal components after applying the gradient boosting model for the training sets from the Ferrenti and empirical approaches are shown in Figure~\ref{fig:PComponents}. 
Further details on the evaluation of the machine learning models and the results from the principal component analyses for all four ML models are shown in the Supplementary Material at \cite{supplementary}.

Fig.~\ref{fig:PComponents} displays the principal components as they are sorted in descending order by the explained variance. For the Ferrenti approach (Fig.~\ref{fig:PComponents}a) to reach the $95 \ \%$ accumulated explained variance, a total of $144$ principal components must be included. Importantly, we observe that the single most important feature for all models in the Ferrenti approach is the fifth principal component. Moreover, by selecting the highest values in this eigenvector, we find that the corresponding features originate from the DFT computed band gap of the elemental solids among the elements in the compound as calculated by Materials Agnostic Platform for Informatics and Exploration (MagPie). The second component's largest values correspond to the 
%electronegativity, ionic property, and covalence radius 
covalent radius, ionic character and packing efficiency 
among the elements in the composition. 
The data originates from elemental calculations from MagPie and are aggregated as either minimum, mean, standard deviation, or maximum. 
While the first principal component encompasses the largest explained variance, it does not provide any specific information on which features it represents.

@@

@@

The gradient boosting model performs optimally at $13$ principal components with a mean F$1$-score of $0.96$. The empirical approach differs in many aspects from the Ferrenti and extended Ferrenti approaches. Firstly, we find that the number of principal components necessary to obtain $95 \ \%$ variance is reduced to $103$ components, which is $41$ and $56$ less than the Ferrenti and extended Ferrenti approaches, respectively. Thus, the variance of the training set is found to be described with fewer principal components, indicating a simpler model.   
Secondly, we find that the first principal component is by far the most important feature for all models in the empirical approach, as visualized in Fig.~\ref{fig:PComponents}(b).  This is part of the reason why we experience a large accuracy for only a single feature (see also the Supplementary Material at \cite{supplementary} fur further details). The first principal component's corresponding features are challenging to explain due to small variations of values, but we find that it includes bond orientational parameters, coordination numbers, and the radial distribution function of a compound's crystal system. 
The standard deviation of the radial distribution function appears multiple times in the list of features and appears to be of particular importance. 
Thirdly, the empirical approach differs in how much explained variation is retained by the first component, which is $21 \ \%$, while it is $14 \ \%$ for the Ferrenti  and $11 \ \%$ for the extended Ferrenti approach. We find the difference striking considering that the approaches share the same ultimate goal, but where the training sets apparently exhibit large and significant variations.

Considering the principal components that are deemed important by the three approaches, we find several intriguing differences. The Ferrenti and extended Ferrenti approaches place particular value on the band gap, of both the predicted compounds and the elemental solids, as well as features such as the ionic or covalent character of the bonding to matter. This is interesting as these features were used as guidance in the initial selection process when forming the training sets. The machine learning models appear to recognize at least some of the expected features as important. Intriguingly, for the empirical approach, other properties dominate. While, e.g., the band gap may still be important, the ML models appear to be informed by other characteristics more related to symmetry and crystal structure, based on the repeated appearance of bond length, orientation and radial distribution in the first principal component. In this case, we did not guide the selection process in terms of specific properties, as was done in the Ferrenti based approaches. Instead, the ML models appear to be inferring that the common trend among the materials that are empirically known to be suitable for QT is related to more complex mechanisms in the crystal structure and bonding.

@@

We first consider the machine learning classification of the test set derived in the Ferrenti approach. The number of candidates predicted by each of the four machine learning methods logistic regression, decision trees, random forests and gradient boosting is contained in Figure~6 of the Supplementary Material at \cite{supplementary} for each of the three approaches. 
%MEB We can take this out, I suppose, if we need more space, and put it in Supplementary (I put fig 5 there) 
From the training set of $23.623$ materials, the four methods consistently predict at least $11.000$ materials as promising candidates in the case of the Ferrenti approach. All four ML methods agree on a total of $6804$ suitable candidates, however, many of these materials are predicted with an incidence similar to that of a coin flip. If we were to raise the minimum confidence cut-off of a prediction to (e.g.) $0.7$ instead of $0.5$, the above ML algorithms would only agree on $1784$ suitable candidates. 
We find that all four ML methods admit almost all materials with a chemical formula matching the known suitable candidates (see  Table~\ref{tab:qt-materials}) that were present in the training set. 
This can allow materials with unfortunate structures to be labeled as suitable candidates by all of our ML methods. 
For example, the ML methods do not perpetuate the band gap restriction from the training set definition. 
Based on the restrictions we defined initially, we would expect that all materials with band gaps lower than $0.5$~eV would be eliminated 
%for a cut-off set to above $0.5$ 
after applying the ML algorithms.  
Although this is expected behavior based on the training set, the trend is not carried over to the predicted data. We find that many entries with band gaps lower than $0.5$~eV are predicted as both suitable and unsuitable candidates by all four ML methods employed herein, 
indicating that the exact cut-off value we set for the band gap ($0.5$~eV for the Ferrenti approach) is not recognized by the machine learning algorithms --- despite the PC analysis revealing the band gap as an important feature for the ML classification in the Ferrenti approach.  
%This cut-off value was imposed on the training data because small band gaps could make it challenging to facilitate deep level defects \cite{Weber2010}. 
On the other hand, due to the known underestimation of band gaps by the PBE functional, the band gaps could in reality be larger for many of the materials.

\subsubsection*{The extended Ferrenti approach}
Next, we turn to the more liberal extended Ferrenti approach containing a $78 \ \%$ larger 
%amount of materials 
data set 
than the Ferrenti approach.  
The four ML methods predict at least $13.000$ materials as suitable candidates out of $22.550$ materials in the training set, where they all agree on a total of $9227$ predicted suitable candidates. Comparing to the training set reveals that all of the unlabeled candidates that are known to be suitable (see Table~\ref{tab:qt-materials}) are, in fact, predicted as suitable candidates. Unfortunately, the ML methods also yield predictions that do not match with what one would expect from physics informed  considerations. 
All four ML algorithms confidently (to confidences of $0.83$ and $0.60$ for two different configurations) predict NaCl as a suitable candidate, which we believe is unlikely due to the strong electrostatic interactions between Na and Cl and the ionic character of their bonding (see, e.g., Ref.~\cite{Weber2010}).  
Indeed, single-photon emitters with sharp and bright zero-phonon lines are uncommon in materials with a strong ionic character. 
Moreover, despite enforcing an even more conservative band gap restriction of $1.5$~eV for the training set as compared to the Ferrenti approach, we still find that all four implemented ML methods predict suitable candidates that exhibit band gaps substantially lower than $0.5$~eV.
Clearly, the ML methods are not adhering rigidly to our imposed band gap restriction when classifying the materials. 
As we based this band gap restriction on assumptions for important material parameters in the QT context, the ML methods might be picking up on something we are not aware of.

\begin{figure}[t]
    \centering
    \includegraphics[width=0.50\textwidth]{figures/vesta-images/cubic-5-structures.png}
    \includegraphics[width=0.23\textwidth]{figures/vesta-images/hex-CdSe-CuI.png}
    \includegraphics[width=0.16\textwidth]{figures/vesta-images/BC2N.png}
    \caption{Illustrations of the crystal structures of the $9$ materials predicted by the empirical approach to $0.85$ confidence. Three different symmetry categories are clearly observable: (a) cubic, (b) hexagonal and (c) \textit{weird} (orthorombic top, chalcopyrite bottom).    }
    \label{fig:vesta-empirical-0p85}
\end{figure}

@@

Consider now the $9$ materials that were classified by the empirical approach as suitable to a confidence of $0.85$ or higher by all four ML models; BN, CdSe, BC$_2$N, InAs, CuI, and ZnCd$_3$Se$_4$. 
%include this? 
The crystal structures of these materials are schematically drawn in Figure~\ref{fig:vesta-empirical-0p85} and grouped into three categories according to whether they are (a) cubic, (b) hexagonal or (c) other. 
%  
The compound BN (mp-$1639$) was in fact present in the training data as a suitable candidate, and we believe this a partial explanation for why the models recognize BN  as a suitable candidate with high probability. Furthermore, two compositions of CdSe (the cubic mp-$2691$ and the hexagonal mp-$1070$) have been predicted as suitable, possibly as a consequence of the similar compound CdS being labeled as a suitable candidate in the training set. 
%The element S resides in the same group as Se, and 
The two compounds of CdSe share similar properties with calculated band gaps in the MP database of $0.5$~eV and $0.6$~eV, respectively.  
Comparing to experiment, ... 
%high quality nanocrystals of CdS have been synthesized and studied for use as quantum dots since the nanocrystals emit in the visible range with a direct band gap of $2.5$-$2.6$~eV \cite{BanerjeeR2000Eots,CelebiSerdar2007SaCo,Qutub2012}. 
%Bulk CdS has a slightly lower reported band gap of $2.4$~eV \cite{Qutub2012}.

In the case of BC$_2$N we find two compositions with the same chemical formula; the orthorhombically coordinated (mp-$629458$) and the chalcopyrite structured (mp-$1008523$) BC$_2$N. The former takes on a polar space group while the latter does not. The band gaps are listed as $1.9$~eV and $1.7$~eV, respectively, in the MP database. 
%BC$_2$N is known as heterodiamond and is a hybrid of diamond and BN. 
Both structures exhibit, as expected, strong covalent character and have been studied for applications as nanostructures for electronic devices \cite{Gao2017}, hydrogen storage \cite{Cai2017} and superhard materials \cite{Li2017, Jiang2020}. Importantly, the diamond-like structure of BC$_3$N was recently predicted as a prominent spin qubit material host \cite{Wang2020SpinQB}. By creating a boron (B) vacancy one can in theory obtain a defect center with similar properties to those found for the NV center in diamond. Whether this is also possible for BC$_2$N remains to be seen. We note that BC$_3$N was not represented in the MP database at the time of downloading and is therefore not included in our data set.

The compounds InAs (cubic, mp-$20305$), CuI (cubic, mp-$22895$ and hexagonal, mp-$569346$) and ZnCd$_3$Se$_4$ (cubic, mp-$1078597$) are listed in the MP database with band gaps of $0.3$, $1.2$, $1.2$ and $1.7$~eV, respectively. 
%, and exhibit comparable properties. 
To the best of our knowledge, ZnCd$_3$Se$_4$ has yet to be synthesized. 
Self-assembled InAs quantum dots are an exciting possible material to use in quantum technology \cite{Liu2018}. 
CuI has recently been synthesized as single-crystal epitaxial films and was shown to exhibit remarkable optoelectronic properties \cite{Ahn2020}.  Interestingly, the material exhibits a large ionic character with more similarities to some of the oxide compounds in the data set than, e.g., the materials that are more covalent in character like Si, SiC and  diamond.  
It seems that ionic character alone is not an obstacle for a material to be quantum compatible if the prediction of CuI proves to be merited. It is unknown at this time whether any potential quantum compatible properties of CuI would originate from deep level defects or nanostructuring (e.g., quantum dots). Note that NaCl was not predicted as suitable for QT in the empirical approach.

>>>>

The most striking difference is found for the empirical approach. The entries labeled as suitable candidates do not exhibit any magnetization, even though the materials crystallize in both polar and non-polar space groups --- contrarily to the two Ferrenti-based approaches. Moreover, we observe that the range of covalent radius and maximum ionic character span a substantially smaller parameter space in the case of the empirical approach than for the two former ones.  
Here, we can begin to deduce some trends, if the findings from the empirical approach are taken as guidance. Where the polarity of the material's space group may be less important than assumed in the Ferrenti based approaches, ionic character and covalent range are distinguished from the other features visualized in Fig.~\ref{fig:parallel-coordinates-approaches}. 
%Identification of the features that promote quantum compatible  characteristics could be crucial for targeted quantum device design based on careful material selection, nanostructuring and defect engineering.

\begin{figure}[t] %!htp
    \centering
    \begin{subfigure}{0.9\textwidth}
        \centering
        \input{figures/parallel-coordinates-plots/01-ferrenti-approach-v2.pgf}
    \end{subfigure}
    \begin{subfigure}{0.9\textwidth}
        \centering
          \input{figures/parallel-coordinates-plots/02-augmented-ferrenti-approach-v2.pgf}
    \end{subfigure}
    \begin{subfigure}{0.9\textwidth}
        \centering
          \input{figures/parallel-coordinates-plots/03-insightful-approach-v2.pgf}
    \end{subfigure}
    %\vspace*{10mm}
    \caption{Parallel coordinate plots for the different approaches. To limit the data cluttering, we have randomly collected up to $250$ entries for each class and made the lines transparent. For the empirical approach, we have used all the $187$ suitable candidates. The axes show total magnetization (mag) taken from the Materials Project (MP) database along with the space group (SG), ionic character (ionic char), covalent range (cov range) as calculated from elemental properties, number of elements (num elements) and energy gap (Eg) extracted from MP.} 
    \label{fig:parallel-coordinates-approaches}
\end{figure}

\begin{figure}%[h] %!tbp
    \centering
    \begin{subfigure}{0.3\textwidth}
        \centering
        \includegraphics[width=1\textwidth]{figures/pca-2d-plots/01-ferrenti-approach-v3.pdf}
    \end{subfigure}%
    \begin{subfigure}{0.3\textwidth}
        \centering
        \includegraphics[width=1\textwidth]{figures/pca-2d-plots/02-augmented-ferrenti-approach-v2.pdf}
    \end{subfigure}
    \begin{subfigure}{0.4\textwidth}
        \centering
        \includegraphics[width=1\textwidth]{figures/pca-2d-plots/03-insightful-approach-v3.pdf}
    \end{subfigure}
    %\vspace*{-95mm}
    \caption{Two-dimensional scatter plots for the three different approaches. We have identified the two eigenvectors corresponding to the two largest eigenvalues of the covariance-matrix, that is, the two most important principal components (PC1 and PC2) of the initial data from the Materials Project query. Then, we have transformed the three training sets resulting from the three approaches and visualized them as scatter plots. Green squares display suitable candidates, along with the black (SiC), blue (diamond) and yellow (Si) symbols in the bottom panel, and red triangles represent unsuitable candidates.}
    \label{fig:2dscatterplotpca}
\end{figure}

To further illustrate the complexity of the training sets, we have identified the two largest eigenvalues of the covariance matrix \cite{Hastie2009} of the complete initial data set from the Materials Project, and transformed the training sets according to the corresponding two eigenvectors, labeled E1 and E2. 
The resulting scatter plots are depicted in Figure~\ref{fig:2dscatterplotpca} for the three approaches, and the training data of each approach is visualized according to the same model. Note that some minor differences between the three panels in Fig.~\ref{fig:2dscatterplotpca} may occur due to the process of centering the data.

Green squares and red triangles represent suitable and unsuitable candidates, respectively. Due to the complexity of reducing the large amount of features down to only two, both suitable and unsuitable candidates for the Ferrenti and the extended Ferrenti approaches are largely overlapping. 
The logic behind categorizing materials in two classes (suitable and unsuitable) appears to not translate into a distinct separation, at least not in the representation of  Fig.~\ref{fig:2dscatterplotpca} for the Ferrenti and extended Ferrenti approaches.  
Hence, using these approaches for predicting QT material hosts could prove challenging for any model that would try to glean a clear-cut boundary between materials that are and are not suitable for QT. 
We expect therefore that the two Ferrenti approaches could need supplementary dimensions for further distinguishing between the materials in the two categories (suitable and unsuitable), or that it may prove challenging to find a generalized model based on the training sets that were defined according to the above discussion.

In the case of the empirical approach, on the other hand, we can discern a trend where the upper left part of the figure is dominated by suitable candidates (green, blue, yellow and black symbols). 
Interestingly, we observe that different configurations of the famously quantum compatible materials of diamond (blue), silicon carbide (black) and silicon (yellow) are grouped together but each material class is separated in its own region. 
The unsuitable candidates (red squares) are similarly restricted to the lower right corner and there appears to be a clear barrier between the two (suitable and unsuitable), albeit with some exceptions. This indicates that the empirical approach may be capable of separating materials based on their underlying properties. Thus, devising a logical framework for the data mining process is likely important in order to obtain a suitable training set for the machine learning methods.

@@

%summary - do we need this? 
In summary, using various data extraction tools for a number of materials science databases, we have obtained a data set consisting of over $25.000$ materials. The information from this data set has then been converted into nearly $5000$ physics-informed features.  
After selecting and fine-tuning the entries for our training sets we are now ready to apply the machine learning methods on the test set in order to classify materials as potential candidates for quantum technology applications.

@@

Four well-tested machine learning algorithms were employed in this work: logistic regression, decision trees, and ensemble methods like random forests and gradient boosting \cite{Mehta2019,Hastie2009}. 
These four machine learning methods were applied to the training sets that were optimized using each of the three approaches (Ferrenti, extended Ferrenti and empirical) outlined above. This yielded twelve sets of materials that were predicted as suitable material hosts for QT applications. 
%
In order to deduce how the different material features are weighted by the four ML models when classifying materials into suitable and unsuitable candidates for QT, we perform a principal component analysis (PCA) for each of the three approaches.   
First, the eigenvalues of the covariance matrices of the three training sets are identified in order of decreasing value,  resulting in three distinct PCA models (as opposed to the case of Fig.~\ref{fig:2dscatterplotpca} where all panels are based on the same model). 
Next, machine learning if performed. 
% 
In the evaluation of the machine learning models for the three different approaches we apply a $5\times 5$ stratified cross-validation when searching for the optimal hyperparameter  combinations. 
%
The extracted principal components after applying the gradient boosting model for the training sets from the Ferrenti and empirical approaches are shown in Figure~\ref{fig:PComponents} in decreasing order of eigenvalue size. 
Further details on the evaluation of the machine learning models and the results from the principal component analyses for all four ML models are shown in the Supplementary Material at \cite{supplementary}.

@@

Fig.~\ref{fig:PComponents} displays the principal components as they are sorted in descending order by the explained variance. For the Ferrenti approach (Fig.~\ref{fig:PComponents}a) to reach the $95 \ \%$ accumulated explained variance, a total of $144$ principal components must be included. Importantly, we observe that the single most important feature for all models in the Ferrenti approach is the fifth principal component (fifth largest eigenvalue). Moreover, by selecting the highest values in this eigenvector, we find that the corresponding features originate from the DFT computed band gap of the elemental solids among the elements in the compound as calculated by Materials Agnostic Platform for Informatics and Exploration (MagPie). 
The second most important PC for the Ferrenti approach is the second one, PC2. 
The second component's largest values correspond to the 
%electronegativity, ionic property, and covalence radius 
covalent radius, ionic character and packing efficiency  
among the elements in the composition. 
The data originates from elemental calculations from MagPie and are aggregated as either minimum, mean, standard deviation, or maximum. 
While the first principal component encompasses the largest explained variance, it does not provide any specific information on which features it represents.

@@

The gradient boosting model performs optimally at $13$ principal components with a mean F$1$-score of $0.96$. 
The empirical approach differs in many aspects from the Ferrenti and extended Ferrenti approaches. 
Firstly, we find that the number of principal components necessary to obtain $95 \ \%$ variance is reduced to $103$ components, which is $41$ and $56$ less than the Ferrenti and extended Ferrenti approaches, respectively. Thus, the variance of the training set is found to be described with fewer principal components, indicating a simpler model.   
Secondly, the first principal component is by far the most important feature for all models in the empirical approach, as visualized in Fig.~\ref{fig:PComponents}(b). 
This is part of the reason why we experience a large accuracy for only a single feature (see also the Supplementary Material at \cite{supplementary} fur further details). The first principal component's corresponding features are challenging to explain due to small variations of values, but we find that it includes bond orientational parameters, coordination numbers, and the radial distribution function of a compound's crystal system. 
The standard deviation of the radial distribution function appears multiple times in the list of features and appears to be of particular importance. 
Thirdly, the empirical approach differs in how much explained variation is retained by the first component, which is $21 \ \%$, while it is $14 \ \%$ for the Ferrenti  and $11 \ \%$ for the extended Ferrenti approach. We find the difference striking considering that the approaches share the same ultimate goal, but where the training sets apparently exhibit large and significant variations.

The principal components that are deemed important by the three approaches are distinct on multiple levels. The Ferrenti and extended Ferrenti approaches place particular value on the band gap, of both the predicted compounds and the elemental solids, as well as features such as the ionic or covalent character of the bonding. Interestingly, these features were used as guidance in the initial selection process when forming the training sets. The machine learning models appear to preserve the criteria imposed in the Ferrenti approaches. Intriguingly, for the empirical approach, other properties dominate. Rather than focusing on, e.g., the band gap, the ML models appear to be informed by other characteristics more related to symmetry and crystal structure, based on the repeated appearance of bond length, orientation and radial distribution in the first principal component. In this case, we did not guide the selection process in terms of specific properties, as was done in the Ferrenti based approaches. Instead, the ML models appear to be inferring that the common trend among the materials that are suitable for QT is related to more complex mechanisms in the crystal structure and bonding.

@@

We first consider the machine learning classification of the test set derived in the Ferrenti approach. The number of candidates predicted by each of the four machine learning methods logistic regression, decision trees, random forests and gradient boosting is visualized in the Supplementary Material at \cite{supplementary} for each of the three approaches. 
%MEB We can take this out, I suppose, if we need more space, and put it in Supplementary (I put fig 5 there) 
From the training set of $23.623$ materials, the four methods consistently predict at least $11.000$ materials as promising candidates in the case of the Ferrenti approach. All four ML methods agree on a total of $6804$ suitable candidates, however, many of these materials are predicted with an incidence similar to that of a coin flip. If we were to raise the minimum confidence cut-off for a prediction to (e.g.) $0.7$ instead of $0.5$, the above ML algorithms would only agree on $1784$ suitable candidates. 
We find that all four ML methods admit almost all materials with a chemical formula matching the known suitable candidates (see  Table~\ref{tab:qt-materials}) that were present in the training set. 
This can allow materials with unfortunate structures to be labeled as suitable candidates by all of our ML methods. 
For example, the ML methods do not perpetuate the band gap restriction from the training set definition. 
Based on the restrictions we defined initially, we would expect that all materials with band gaps lower than $0.5$~eV would be eliminated 
%for a cut-off set to above $0.5$ 
after applying the ML algorithms.  
Although this is expected behavior based on the training set, the trend is not carried over to the predicted data. We find that many entries with band gaps lower than $0.5$~eV are predicted as both suitable and unsuitable candidates by all four ML methods employed herein, 
indicating that the exact cut-off value we set for the band gap ($0.5$~eV for the Ferrenti approach) is not recognized by the machine learning algorithms --- despite the PC analysis revealing the band gap as an important feature for the ML classification in the Ferrenti approach.  
On the other hand, due to the known underestimation of band gaps by the PBE functional, the band gaps could in reality be larger for many of the materials.

\subsubsection*{The extended Ferrenti approach}
Next, we turn to the more liberal extended Ferrenti approach containing a $78 \ \%$ larger 
%amount of materials 
data set 
than the Ferrenti approach.  
The four ML methods predict at least $13.000$ materials as suitable candidates out of $22.550$ materials in the training set, where they all agree on a total of $9227$ predicted suitable candidates. Comparing to the training set reveals that all of the unlabeled candidates that are known to be suitable (see Table~\ref{tab:qt-materials}) are, in fact, predicted as suitable candidates. 
% rewrite?? 
However, the ML methods also yield predictions that do not match with the expectation that we are looking for a covalently bonded semiconductor to host a sharp and bright quantum emitter. 
% 
All four ML algorithms confidently (to confidences of $0.83$ and $0.60$ for two different configurations) predict NaCl as a suitable candidate, which was unexpected due to the strong electrostatic interactions between Na and Cl and the ionic character of their bonding (see, e.g., Ref.~\cite{Weber2010}).  
Indeed, single-photon emitters with sharp and bright zero-phonon lines are uncommon in materials with a strong ionic character. 
Moreover, despite enforcing an even more conservative band gap restriction of $1.5$~eV for the training set as compared to the Ferrenti approach, we still find that all four implemented ML methods predict suitable candidates that exhibit band gaps substantially lower than $0.5$~eV.
Clearly, the ML methods are not adhering rigidly to our imposed band gap restriction when classifying the materials. 
As we based this band gap restriction on assumptions for important material parameters in the QT context, the ML methods might be picking up on something we did not anticipate.

@@

\begin{figure}[t]
    \centering
    \includegraphics[width=0.85\textwidth]{figures/vesta-images/fig6-all.png}
    \caption{Illustrations of the crystal systems that the $9$ materials predicted by the empirical approach to $0.85$ confidence, and the $6$ materials predicted by all approaches to $0.75$ confidence, may belong to.  Four different symmetry categories are observed; cubic, hexagonal, tetragonal and orthorombic.   }
    \label{fig:crystalsystems}
\end{figure}

\begin{table}[b]
    \centering 
    \caption{Grouping of predicted materials (above $0.75$ confidence) into the four appearing crystal systems. The $9$ materials predicted by the empirical approach ($>0.85$) and the $6$ materials predicted by all approaches ($>0.75$) are considered. }
    \begin{tabular}{c|c c c c}
      & Cubic & Hexagonal & Tetragonal & Orthorombic \\
    \hline 
    Empirical approach & CuI (mp-$22895$) & CuI (mp-$569346$) & BC$_2$N (mp-$1008523$) & BC$_2$N (mp-$629458$) \\
    to $0.85$ confidence & CdSe (mp-$2691$) & CdSe (mp-$1070$) & & \\
     & BN (mp-$1639$) & & & \\
     & InAs (mp-$20305$) & & & \\
     & ZnCd$_3$Se$_4$ (mp-$1078597$) & & & \\
    \hline 
    All approaches & InP (mp-$20351$) & InP (mp-$966800$) & ZnGeAs$_2$ (mp-$4008$) & \\ 
    to $0.75$ confidence & GeSe (mp-$10759$) & & CdSnP$_2$ (mp-$5213$) & \\ 
     & SiSn (mp-$1009813$) & & & \\
    \end{tabular}
    \label{tab:crystalsystems}
\end{table}

Consider now the $9$ materials that were classified by the empirical approach as suitable to a confidence of $0.85$ or higher by all four ML models; BN, CdSe, BC$_2$N, InAs, CuI, and ZnCd$_3$Se$_4$. 
The nine materials each belong to one of the four crystal systems cubic, hexagonal, tetragonal and orthorombic.  Figure~\ref{fig:vesta-empirical-0p85} visualizes the four different crystal systems by means of examples and Table~\ref{tab:crystalsystems} categorizes each material according to its point group as listed in the MP database. 
Interestingly, we find that all nine materials are four-fold coordinated, and that the first two lattice vector are identical in all nine cases whereas $c$ may differ.  
There appears to be a high degree of symmetry present for all materials but we find no elemental and perfectly symmetric semiconductors in the list.

Turning to each of the individual material predictions, the compound BN (mp-$1639$) was in fact present in the training data as a suitable candidate, and we believe this a partial explanation for why the models recognize BN as a suitable candidate with high probability. Furthermore, two compositions of CdSe (the cubic mp-$2691$ and the hexagonal mp-$1070$) have been predicted as suitable, possibly as a consequence of the similar compound CdS being labeled as a suitable candidate in the training set. 
%The element S resides in the same group as Se, and 
The two compounds of CdSe share similar properties with calculated band gaps in the MP database of $0.5$~eV and $0.6$~eV, respectively. Comparing to experiment, ... 
%high quality nanocrystals of CdS have been synthesized and studied for use as quantum dots since the nanocrystals emit in the visible range with a direct band gap of $2.5$-$2.6$~eV \cite{BanerjeeR2000Eots,CelebiSerdar2007SaCo,Qutub2012}. 
%Bulk CdS has a slightly lower reported band gap of $2.4$~eV \cite{Qutub2012}.

In the case of BC$_2$N we find two compositions with the same chemical formula; the orthorhombically coordinated (mp-$629458$) and the tetragonally structured (mp-$1008523$) BC$_2$N. The former takes on a polar space group while the latter does not. The band gaps are listed as $1.9$~eV and $1.7$~eV, respectively, in the MP database. 
%BC$_2$N is known as heterodiamond and is a hybrid of diamond and BN. 
Both structures exhibit, as expected, strong covalent character and have been studied for applications as nanostructures for electronic devices \cite{Gao2017}, hydrogen storage \cite{Cai2017} and superhard materials \cite{Li2017, Jiang2020}. Importantly, the diamond-like structure of BC$_3$N was recently predicted as a prominent spin qubit material host \cite{Wang2020SpinQB}. By creating a boron (B) vacancy one can in theory obtain a defect center with similar properties to those found for the NV center in diamond. Whether this is also possible for BC$_2$N remains to be seen. We note that BC$_3$N was not represented in the MP database at the time of downloading and is therefore not included in our data set.

The compounds InAs (cubic, mp-$20305$), CuI (cubic, mp-$22895$ and hexagonal, mp-$569346$) and ZnCd$_3$Se$_4$ (cubic, mp-$1078597$) are listed in the MP database with band gaps of $0.3$, $1.2$, $1.2$ and $1.7$~eV, respectively. 
%, and exhibit comparable properties. 
To the best of our knowledge, ZnCd$_3$Se$_4$ has yet to be synthesized. 
Self-assembled InAs quantum dots are an exciting possible material to use in quantum technology \cite{Liu2018}. 
CuI has recently been synthesized as single-crystal epitaxial films and was shown to exhibit remarkable optoelectronic properties \cite{Ahn2020}. Interestingly, the material exhibits a large ionic character with more similarities to some of the oxide compounds in the data set than, e.g., the materials that are more covalent in character like Si, SiC and  diamond.  
It seems that ionic character alone is not an obstacle for a material to be quantum compatible if the prediction of CuI proves to be merited. It is unknown at this time whether any potential quantum compatible properties of CuI would originate from deep level defects or nanostructuring (e.g., quantum dots). NaCl was not predicted as suitable for QT in the empirical approach.
